---
title: "Summary metrics for paper"
output: html_document
---

This file contains the summary metrics reported in the paper in one place.
```{r, echo = FALSE}
library(targets)
library(dplyr)
library(kableExtra)
library(scoringutils)
library(tidyr)

# delays
clean_weekly_data <- tar_read(clean_weekly_data)

# scores
scores_su <- tar_read(scores_su)
scores_su_raw_median <- tar_read(scores_su_raw_median)
scores_ag_su_raw_median <- tar_read(scores_ag_su_raw_median)
scores_ag_su <- tar_read(scores_ag_su)
scores_ag_su_all <- tar_read(scores_ag_su_all)

# training volume allocation
map_tv <- tar_read(map_tv)
```

# Delays
```{r}
# Average delay time by pathogen in 2024-2025
clean_weekly_data |>
  group_by(pathogen_name, season) |>
  summarise(mean_delay = sum(count * delay) / sum(count)) |>
  mutate(age_group = "00+") |>
  filter(season == "2024-2025") |>
  kable(caption = "Pathogen specific mean delays 2024-2025")
```
# Table of proportion complete by horizon week and pathogen
```{r}
final_data <- clean_weekly_data |>
  group_by(pathogen, end_of_week_reference_date) |>
  summarise(final_count = sum(count))

full_prop_reported_table <- clean_weekly_data |>
  group_by(
    pathogen, end_of_week_reference_date,
    end_of_week_report_date, delay, season
  ) |>
  summarise(count = sum(count)) |>
  group_by(pathogen, end_of_week_reference_date, season) |>
  mutate(cum_count = cumsum(count)) |>
  left_join(final_data) |>
  mutate(prop_reported = cum_count / final_count) |>
  group_by(pathogen, season, delay) |>
  summarise(
    median_prop_reported = median(prop_reported),
    ub_prop_reported = quantile(prop_reported, 0.975),
    lb_prop_reported = quantile(prop_reported, 0.025)
  )

full_prop_reported_table |>
  filter(delay <= 3, season != "other") |>
  mutate(prop_reported = glue("{round(median_prop_reported, 3)} ({round(lb_prop_reported,3)}, {round(ub_prop_reported,3)})")) |>
  select(pathogen, season, prop_reported, delay) |>
  pivot_wider(names_from = delay, values_from = prop_reported, names_prefix = "delay: ") |>
  kable(caption = "Proportion reported relative to final data by season and pathogen")
```

# Shortest and longest mean delay across age groups in BAR
```{r}
clean_weekly_data |>
  filter(pathogen == "bar") |>
  group_by(pathogen_name, season, age_group) |>
  summarise(mean_delay = 7 * sum(count * delay) / sum(count)) |>
  filter(season == "2024-2025") |>
  kable(caption = "BAR age-specific mean delays 2024-2025")
```

# RSV delays
```{r}
clean_weekly_data |>
  filter(pathogen == "rsv") |>
  group_by(pathogen_name, season, age_group) |>
  summarise(mean_delay = 7 * sum(count * delay) / sum(count)) |>
  filter(season == "2024-2025") |>
  kable(caption = "RSV age-specific mean delays 2024-2025")
```

# State-level nowcast results

```{r}
# rWIS (MADPH/baselinenowcast) for BAR, COVID, flu, RSV
scores_su |>
  summarise_scores(by = c("model", "pathogen")) |>
  select(pathogen, wis, model) |>
  pivot_wider(
    names_from = "model",
    values_from = "wis"
  ) |>
  mutate(rwis = `MADPH method` / baselinenowcast) |>
  kable(caption = "rWIS vs baselinenowcast")
```

## MAE compared to preliminary data
```{r}
scores_su_raw_median |>
  filter(
    scale == "natural",
    model %in% c(
      "baselinenowcast",
      "MADPH method",
      "preliminary data"
    )
  ) |>
  summarise_scores(by = c("model", "pathogen")) |>
  select(model, pathogen, ae_median) |>
  pivot_wider(
    names_from = "model",
    values_from = "ae_median"
  ) |>
  kable(caption = "Median absolute error overall")
```

# Age-group nowcast results

```{r}
# rWIS(baselinenowcast strata sharing/baselinenowcast base) for BAR, COVID, flu, RSV

scores_ag_su |>
  summarise_scores(by = c("model", "pathogen")) |>
  select(pathogen, wis, model) |>
  pivot_wider(
    names_from = "model",
    values_from = "wis"
  ) |>
  mutate(
    rwis_strata = `baselinenowcast strata sharing` / baselinenowcast,
    rwis_madph = `MADPH method` / baselinenowcast
  ) |>
  kable(caption = "rWIS vs baselinenowcast age groups")
```

# MAE by age group
```{r}
scores_ag_su_raw |>
  filter(
    scale == "natural",
    model %in% c(
      "baselinenowcast",
      "MADPH method",
      "preliminary data"
    )
  ) |>
  summarise_scores(by = c("model", "age_group", "pathogen")) |>
  select(model, pathogen, age_group, ae_median) |>
  pivot_wider(
    names_from = "model",
    values_from = "ae_median"
  ) |>
  kable(caption = "Median absolute error overall")
```

# 00-04  & 65+  rWIS for RSV
```{r}
scores_ag_su |>
  summarise_scores(by = c("model", "pathogen", "age_group")) |>
  select(pathogen, wis, model, age_group) |>
  filter(pathogen == "rsv", age_group %in% c("00-04", "65+")) |>
  pivot_wider(
    names_from = "model",
    values_from = "wis"
  ) |>
  mutate(
    rwis_strata = `baselinenowcast strata sharing` / baselinenowcast,
    rwis_madph = `MADPH method` / baselinenowcast
  ) |>
  kable(caption = "rWIS vs baselinenowcast 00-04 & 65+")
```

Comparing methods of producing weekly nowcasts
```{r}
scores_ag_su_all |>
  summarise_scores(by = c("model", "pathogen")) |>
  select(pathogen, wis, model) |>
  pivot_wider(
    names_from = "model",
    values_from = "wis"
  ) |>
  mutate(
    rwis_strata = `baselinenowcast strata sharing` / baselinenowcast,
    rwis_madph = `MADPH method` / baselinenowcast
  ) |>
  kable(caption = "rWIS comparing baselinenowcast methods of producing weekly nowcasts")
```
